---
title: "Untitled"
output: html_document
date: "2025-09-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}

library(here)
library(tidyverse)

```

```{r}


# Filter and calculate percentages
results <- gender_age %>%
  # Filter for England only (exclude Wales)
  filter(Regions != "Wales") %>%
  # Filter for ages 16-24 only (exclude 15 and under, and 25+)
  filter(`Age (B) (11 categories)` %in% c(
    "Aged 16 to 17 years", 
    "Aged 18 to 19 years",
    "Aged 20 to 21 years", 
    "Aged 22 to 24 years"
  )) %>%
  # Create combined age categories
  mutate(
    age_group = case_when(
      `Age (B) (11 categories)` == "Aged 16 to 17 years" ~ "16-17",
      `Age (B) (11 categories)` %in% c("Aged 18 to 19 years", "Aged 20 to 21 years") ~ "18-21",
      `Age (B) (11 categories)` == "Aged 22 to 24 years" ~ "22-24",
      TRUE ~ "Other"
    )
  ) %>%
  # Group by sex and combined age groups
  group_by(`Sex (2 categories)`, age_group) %>%
  summarise(
    age_gender_total = sum(Observation, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Calculate denominators (total by sex) and percentages
  group_by(`Sex (2 categories)`) %>%
  mutate(
    sex_total = sum(age_gender_total),
    percentage = (age_gender_total / sex_total) * 100
  ) %>%
  ungroup()

# View the results
print(results)

# Alternative view - pivot wider for easier comparison
results_wide <- results %>%
  select(`Sex (2 categories)`, age_group, percentage) %>%
  pivot_wider(
    names_from = `Sex (2 categories)`,
    values_from = percentage,
    names_prefix = "percent_"
  )

print(results_wide)

# Summary by sex
sex_summary <- results %>%
  group_by(`Sex (2 categories)`) %>%
  summarise(
    total_count = sum(age_gender_total),
    .groups = "drop"
  )

```

```{r}

gender_age <- read_csv(here("data", "data", "age_gender_ethnicity.csv"))

# Filter and calculate ethnicity percentages
ethnicity_results <- gender_age %>%
  # Filter for England only (exclude Wales)
  filter(Regions != "Wales") %>%
  # Filter for ages 16-24 only
  filter(`Age (B) (11 categories)` %in% c(
    "Aged 16 to 17 years", 
    "Aged 18 to 19 years",
    "Aged 20 to 21 years", 
    "Aged 22 to 24 years"
  )) %>%
  # Create broad ethnicity categories
  mutate(
    broad_ethnicity = case_when(
      str_detect(`Ethnic group (20 categories)`, "Asian") ~ "Asian",
      str_detect(`Ethnic group (20 categories)`, "Black") ~ "Black", 
      str_detect(`Ethnic group (20 categories)`, "White") ~ "White",
      str_detect(`Ethnic group (20 categories)`, "Mixed") ~ "Mixed",
      TRUE ~ "Other"
    )
  ) %>%
  # Group by broad ethnicity categories
  group_by(broad_ethnicity) %>%
  summarise(
    ethnicity_total = sum(Observation, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Calculate percentages
  mutate(
    total_population = sum(ethnicity_total),
    percentage = (ethnicity_total / total_population) * 100
  ) %>%
  # Arrange by percentage (largest first)
  arrange(desc(percentage))

# View the results
print(ethnicity_results)

# Create a summary table
print("Ethnicity Distribution (16-24 years, England):")
ethnicity_summary <- ethnicity_results %>%
  select(broad_ethnicity, ethnicity_total, percentage) %>%
  mutate(percentage = round(percentage, 2))

print(ethnicity_summary)

# Check what specific ethnicity categories were in the original data
print("Original ethnicity categories in your data:")
unique_ethnicities <- gender_age %>%
  filter(Regions != "Wales") %>%
  filter(`Age (B) (11 categories)` %in% c(
    "Aged 16 to 17 years", 
    "Aged 18 to 19 years",
    "Aged 20 to 21 years", 
    "Aged 22 to 24 years"
  )) %>%
  distinct(`Ethnic group (20 categories)`) %>%
  arrange(`Ethnic group (20 categories)`)

```

```{r}

gender_age <- read_csv(here("data", "data", "age_gender_ethnicity.csv"))

# Filter and calculate regional percentages
region_results <- gender_age %>%
  # Filter out Wales
  filter(Regions != "Wales") %>%
  # Filter for ages 16-24 only
  filter(`Age (B) (11 categories)` %in% c(
    "Aged 16 to 17 years", 
    "Aged 18 to 19 years",
    "Aged 20 to 21 years", 
    "Aged 22 to 24 years"
  )) %>%
  # Group by regions
  group_by(Regions) %>%
  summarise(
    region_total = sum(Observation, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Calculate percentages
  mutate(
    total_population = sum(region_total),
    percentage = (region_total / total_population) * 100
  ) %>%
  # Arrange by percentage (largest first)
  arrange(desc(percentage))

# View the results
print(region_results)


print(unique_ethnicities)

```

``{r}
# Filter and calculate gender percentages
gender_results <- gender_age %>%
  # Filter out Wales
  filter(Regions != "Wales") %>%
  # Filter for ages 16-24 only
  filter(`Age (B) (11 categories)` %in% c(
    "Aged 16 to 17 years", 
    "Aged 18 to 19 years",
    "Aged 20 to 21 years", 
    "Aged 22 to 24 years"
  )) %>%
  # Group by gender
  group_by(`Sex (2 categories)`) %>%
  summarise(
    gender_total = sum(Observation, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Calculate percentages
  mutate(
    total_population = sum(gender_total),
    percentage = (gender_total / total_population) * 100
  ) %>%
  # Arrange by percentage (largest first)
  arrange(desc(percentage))

# View the results
print(gender_results)

```

```{r}


gender_age <- read_csv(here("data", "data", "age_gender_ethnicity.csv"))


library(tidyverse)
library(here)

# Read the data
gender_age <- read_csv(here("data", "data", "age_gender_ethnicity.csv"))

# Filter and calculate percentages with total population as denominator
results <- gender_age %>%
  # Filter for England only (exclude Wales)
  filter(Regions != "Wales") %>%
  # Filter for ages 16-24 only - INCLUDING 16-17!
  filter(`Age (B) (11 categories)` %in% c(
    "Aged 16 to 17 years",
    "Aged 18 to 19 years",
    "Aged 20 to 21 years", 
    "Aged 22 to 24 years"
  )) %>%
  # Create combined age categories - INCLUDING 16-17!
  mutate(
    age_group = case_when(
      `Age (B) (11 categories)` == "Aged 16 to 17 years" ~ "16-17",
      `Age (B) (11 categories)` %in% c("Aged 18 to 19 years", "Aged 20 to 21 years") ~ "18-21",
      `Age (B) (11 categories)` == "Aged 22 to 24 years" ~ "22-24",
      TRUE ~ "Other"
    )
  ) %>%
  # Group by sex and combined age groups
  group_by(`Sex (2 categories)`, age_group) %>%
  summarise(
    age_gender_total = sum(Observation, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Calculate total population across ALL age-gender combinations
  mutate(
    total_population = sum(age_gender_total),
    percentage = (age_gender_total / total_population) * 100
  ) %>%
  # Arrange by sex and age group for easy viewing
  arrange(`Sex (2 categories)`, age_group)

# View the results
print(results)

# Check that percentages sum to 100
print(paste("Total percentage check:", round(sum(results$percentage), 2), "%"))

# Alternative view - pivot wider for easier comparison
results_wide <- results %>%
  select(`Sex (2 categories)`, age_group, percentage) %>%
  pivot_wider(
    names_from = `Sex (2 categories)`,
    values_from = percentage,
    names_prefix = "percent_"
  )

print("Age-Gender Distribution (16-24 years, England):")
print(results_wide)

# Summary by sex (showing what % of total population each sex represents)
sex_summary <- results %>%
  group_by(`Sex (2 categories)`) %>%
  summarise(
    total_count = sum(age_gender_total),
    total_percentage = sum(percentage),
    .groups = "drop"
  )

print("Summary by Sex (% of total 16-24 population):")
print(sex_summary)