---
title: "script"
output: html_document
date: "2025-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}

library(tidyverse)
library(here)

```

```{r}

gender_age <- read_csv(here("data", "gender_age_ew.csv"))

```

```{r}




# Filter to people with young children (under 11) and relevant age groups
gender_age_small <- gender_age %>% 
  filter(`Dependent children in household and their age (6 categories)` %in% 
         c("Youngest dependent child: Aged 0 to 4 years", 
           "Youngest dependent child: Aged 5 to 11 years") &
         `Age (14 categories)` %in% 
         c("Aged 18 to 19 years", "Aged 20 to 21 years", "Aged 22 to 24 years", 
           "Aged 25 to 29 years", "Aged 30 to 34 years", "Aged 35 to 39 years",    
           "Aged 40 to 44 years", "Aged 45 to 49 years", "Aged 50 to 54 years"))

# Create new age groups (18-24, 25-34, 35-44, 45-54)
gender_age_grouped <- gender_age_small %>%
  mutate(
    age_group = case_when(
      `Age (14 categories)` %in% c("Aged 18 to 19 years", "Aged 20 to 21 years", "Aged 22 to 24 years") ~ "18-24",
      `Age (14 categories)` %in% c("Aged 25 to 29 years", "Aged 30 to 34 years") ~ "25-34", 
      `Age (14 categories)` %in% c("Aged 35 to 39 years", "Aged 40 to 44 years") ~ "35-44",
      `Age (14 categories)` %in% c("Aged 45 to 49 years", "Aged 50 to 54 years") ~ "45-54"
    )
  ) %>%
  # Sum observations by age group and sex (across all regions)
  group_by(age_group, `Sex (2 categories)`) %>%
  summarise(total_obs = sum(Observation), .groups = "drop")

# Calculate proportions for males with young children
male_age_interlock <- gender_age_grouped %>%
  filter(`Sex (2 categories)` == "Male") %>%
  mutate(
    total_males_with_young_children = sum(total_obs),
    proportion = total_obs / total_males_with_young_children,
    percentage = round(proportion * 100, 1)
  ) %>%
  select(age_group, count = total_obs, total_males_with_young_children, 
         proportion, percentage) %>%
  arrange(age_group)

# Calculate proportions for females with young children  
female_age_interlock <- gender_age_grouped %>%
  filter(`Sex (2 categories)` == "Female") %>%
  mutate(
    total_females_with_young_children = sum(total_obs),
    proportion = total_obs / total_females_with_young_children,
    percentage = round(proportion * 100, 1)
  ) %>%
  select(age_group, count = total_obs, total_females_with_young_children, 
         proportion, percentage) %>%
  arrange(age_group)

# Display results
print("MALE AGE DISTRIBUTION (among men with children under 11):")
print(male_age_interlock)

print("\nFEMALE AGE DISTRIBUTION (among women with children under 11):")
print(female_age_interlock)

```

```{r}

ethnicity <- read_csv(here("data", "ethnicity_ew.csv"))



# Filter to people with young children (under 11) and relevant age groups (18-54)
ethnicity_filtered <- ethnicity %>% 
  filter(`Dependent children in household and their age (6 categories)` %in% 
         c("Youngest dependent child: Aged 0 to 4 years", 
           "Youngest dependent child: Aged 5 to 11 years") &
         `Age (14 categories)` %in% 
         c("Aged 18 to 19 years", "Aged 20 to 21 years", "Aged 22 to 24 years", 
           "Aged 25 to 29 years", "Aged 30 to 34 years", "Aged 35 to 39 years",    
           "Aged 40 to 44 years", "Aged 45 to 49 years", "Aged 50 to 54 years") &
         `Ethnic group (6 categories)` != "Does not apply")  # Exclude "Does not apply"

# Sum observations by ethnic group (across all regions, ages, and genders)
ethnicity_grouped <- ethnicity_filtered %>%
  group_by(`Ethnic group (6 categories)`) %>%
  summarise(total_obs = sum(Observation), .groups = "drop")

# Calculate proportions for each ethnic group
ethnicity_interlock <- ethnicity_grouped %>%
  mutate(
    total_parents_with_young_children = sum(total_obs),
    proportion = total_obs / total_parents_with_young_children,
    percentage = round(proportion * 100, 1)
  ) %>%
  select(ethnic_group = `Ethnic group (6 categories)`, 
         count = total_obs, 
         total_parents_with_young_children, 
         proportion, 
         percentage) %>%
  arrange(desc(percentage))

# Display results
print("ETHNIC GROUP DISTRIBUTION (among all parents aged 18-54 with children under 11):")
print(ethnicity_interlock)

# Create a cleaner summary table
ethnicity_summary <- ethnicity_interlock %>%
  select(ethnic_group, count, percentage) %>%
  mutate(
    ethnic_group = case_when(
      ethnic_group == "Asian, Asian British or Asian Welsh" ~ "Asian",
      ethnic_group == "Black, Black British, Black Welsh, Caribbean or African" ~ "Black",
      ethnic_group == "Mixed or Multiple ethnic groups" ~ "Mixed",
      ethnic_group == "Other ethnic group" ~ "Other",
      TRUE ~ ethnic_group
    )
  )

print("\nSUMMARY TABLE:")
print(ethnicity_summary)

# Verify percentages add up to 100%
total_percentage <- sum(ethnicity_interlock$percentage)
print(paste("\nTotal percentage check:", round(total_percentage, 1), "%"))


```

```{r}

gender_age <- read.csv("gender_age_ew.csv")




# Filter to people with young children (under 11) and relevant age groups (18-54)
regional_filtered <- gender_age %>% 
  filter(`Dependent children in household and their age (6 categories)` %in% 
         c("Youngest dependent child: Aged 0 to 4 years", 
           "Youngest dependent child: Aged 5 to 11 years") &
         `Age (14 categories)` %in% 
         c("Aged 18 to 19 years", "Aged 20 to 21 years", "Aged 22 to 24 years", 
           "Aged 25 to 29 years", "Aged 30 to 34 years", "Aged 35 to 39 years",    
           "Aged 40 to 44 years", "Aged 45 to 49 years", "Aged 50 to 54 years"))

# Sum observations by region (across all ages and genders)
regional_grouped <- regional_filtered %>%
  group_by(Regions) %>%
  summarise(total_obs = sum(Observation), .groups = "drop")

# Calculate proportions for each region
regional_interlock <- regional_grouped %>%
  mutate(
    total_parents_with_young_children = sum(total_obs),
    proportion = total_obs / total_parents_with_young_children,
    percentage = round(proportion * 100, 1)
  ) %>%
  select(region = Regions, 
         count = total_obs, 
         total_parents_with_young_children, 
         proportion, 
         percentage) %>%
  arrange(desc(percentage))

# Display results
print("REGIONAL DISTRIBUTION (among all parents aged 18-54 with children under 11):")
print(regional_interlock)


regional_summary <- regional_interlock %>%
  select(region, count, percentage)

print("\nSUMMARY TABLE:")
print(regional_summary)


total_percentage <- sum(regional_interlock$percentage)
print(paste("\nTotal percentage check:", round(total_percentage, 1), "%"))

# Optional: Create a more detailed breakdown showing the actual numbers
print("\nDETAILED BREAKDOWN:")
detailed_breakdown <- regional_interlock %>%
  mutate(
    population_thousands = round(count / 1000, 1)
  ) %>%
  select(region, population_thousands, percentage) %>%
  arrange(desc(percentage))

print(detailed_breakdown)

```

```{r}


education <- read_csv(here("data", "education.csv"))



# Read the data (assuming it's called education)
# education <- read.csv("education.csv")

# Filter to people with young children (under 11) and relevant age groups (18-54)
education_filtered <- education %>% 
  filter(`Dependent children in household and their age (6 categories)` %in% 
         c("Youngest dependent child: Aged 0 to 4 years", 
           "Youngest dependent child: Aged 5 to 11 years") &
         `Age (14 categories)` %in% 
         c("Aged 18 to 19 years", "Aged 20 to 21 years", "Aged 22 to 24 years", 
           "Aged 25 to 29 years", "Aged 30 to 34 years", "Aged 35 to 39 years",    
           "Aged 40 to 44 years", "Aged 45 to 49 years", "Aged 50 to 54 years") &
         `Highest level of qualification (7 categories)` != "Does not apply")  # Exclude "Does not apply"

# Create binary degree variable
education_binary <- education_filtered %>%
  mutate(
    degree_status = case_when(
      `Highest level of qualification (7 categories)` == "Level 4 qualifications or above: degree (BA, BSc), higher degree (MA, PhD, PGCE), NVQ level 4 to 5, HNC, HND, RSA Higher Diploma, BTEC Higher level, professional qualifications (for example, teaching, nursing, accountancy)" ~ "Degree",
      TRUE ~ "No degree"
    )
  )

# Sum observations by degree status (across all regions, ages, and genders)
education_grouped <- education_binary %>%
  group_by(degree_status) %>%
  summarise(total_obs = sum(Observation), .groups = "drop")

# Calculate proportions for each education category
education_interlock <- education_grouped %>%
  mutate(
    total_parents_with_young_children = sum(total_obs),
    proportion = total_obs / total_parents_with_young_children,
    percentage = round(proportion * 100, 1)
  ) %>%
  select(degree_status, 
         count = total_obs, 
         total_parents_with_young_children, 
         proportion, 
         percentage) %>%
  arrange(desc(percentage))

# Display results
print("EDUCATION DISTRIBUTION (among all parents aged 18-54 with children under 11):")
print(education_interlock)
